---
title: "Safe data frame subsetting"
author: "GAB"
date: "3/21/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Create a random data frame:

```{r }
df <- data.frame(column1 = 1:10,
                 column2 = 1:10,
                 column3 = 1:10,
                 column4 = 1:10)
print(df)
print(is.data.frame(df))
```

TASK1: subset the first 5 rows of column2
and column 4

The typical way to do it by indexing
```{r }

df_5rows = df[1:5, c(2,4)]
print(df_5rows)
print(is.data.frame(df_5rows))
```
df_5rows is a data.frame, no problem

Task2: subset the first 5 rows of only column1

OPTION #1:

```{r }

df_column1a <- df[1:5, c(1)]
print(df_column1a)
print(is.data.frame(df_column1a))

```

df_column1a is a vector, NOT a data.frame

ok maybe we can convert back to data.frame
```{r }

convert_df <- as.data.frame(df_column1a)
print(convert_df)
print(is.data.frame(convert_df))

```

but convert_df loses original column name and
we need one additional line of code to set it:

```{r }

colnames(convert_df) <- c("column1")
print(convert_df)
print(is.data.frame(convert_df))

```


OPTION #2 (better):

```{r }

df_column1b <- df[1:5, c(1), drop = FALSE]
print(df_column1b)
print(is.data.frame(df_column1b))

```

df_column1b is a data.frame, no problem

NOTES

I) The "matrix" function has this problem as well,
so be sure to use drop = FALSE when subsetting a
matrix

II) data.table does not have this problem

III) data.table is also faster than data.frame so
perhaps use data.table instead of data.frame when
working with data
