---
title: "Ellipsis"
author: "GAB"
date: "8/09/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Define a function with an arbitraty number of arguments
```{r }

fxn_without_named_arguments = function(...){
  return(paste(..., " thank you."))
}
print(fxn_without_named_arguments("my ", "name ", "is ", "R"))

```

Define a function without specifying argument names
```{r }
fxn_with_many_named_arguments = function(...){
  argss <- list(...)
  print(names(argss))
  return(argss$a+argss$b)
}
print(fxn_with_many_named_arguments(a=4, b=9))

```

Using ... can make your function very flexible, but to understand how to use it, users need to read that function's documentation - not just the function's definition - since many arguments will be hidden from the 
function's definition and be replaced with ...

Which means the next time you see a function with ... be sure to read that function's documentation closely if you
want to find out about more arguments you could use.
