---
title: "20200411-joins_example-DR"
author: "DR"
date: "4/11/2020"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Hello R/Python Clubers! I was asked to share an R tip this week. After lots of mulling over and wrestling with whether my tips were common knowledge I decided to post about the elegance and flexibility of joins (which I only recently really came to appreciate). 

https://dplyr.tidyverse.org/reference/join.html

TL;DR: Joins are a set of functions within the dplyr package that allow you to merge and compare data frames in various configurations. They can even streamline logical processes.

 

I've been working on a project lately where I wrote an initial/base script for the data and then had to make versions of that script to fit variations in the initial data formats. When I initially wrote the script I just wanted to get it working correctly, so I ended up with a few messy loops to accomplish the logic needed. I started with something like:

for (i in 1:length(x$column)) {
if (x$column[i,] == “Category 1”)) {
                       x$othercolumn[i,] <- “tag for category 1”
           }else{
                       if (x$column2[i,] == “Category 2”) {
                                   x$othercolumn[i,] <- “tag for #category 2”
           }
}

This worked, but obviously isn’t ideal and would take up a lot of processing capacity if the data were long or there were a lot of categories.

So, on my next iteration I replaced it with:

tagged_data <- left_join(x, tags, by = “id”)

Much cleaner and more efficient!

This is a pretty simple example, but can be applied really broadly and iteratively for complex data sets as needed. (It was also just a good reminder for me to take a step back after a first draft to re-evaluate logic and optimization.)

I also wrote up a quick data set that showcases a couple of the joins I like the most!

load the dplyr library

```{r echo= TRUE, results='hide'}
library(dplyr)
```

initialize some dummy data

```{r }
num <- 1:8
id1 <- as.factor(c("a", "b", "b", "c", "c", "c", "y", "z"))

id2 <- as.factor(c("a", "b", "c", "d", "e"))
col <- as.factor(c("red", "green", "blue", "yellow", "white"))

df1 <- cbind.data.frame(num, id1)
df2 <- cbind.data.frame(id2, col)

print(df1)
print(df2)
```

name the columns so the two data frames have an overlapping variable

```{r }
names(df1) <- c("num", "id")
names(df2) <- c("id", "col")

print(df1)
print(df2)

```

join the data frames in a few different examples

```{r }
merged_left <- left_join(df1, df2, by = "id")
print(merged_left)
```

```{r }
merged_right <- right_join(df1, df2, by = "id")
print(merged_right)
```

```{r }
merged_inner <- inner_join(df1, df2, by = "id")
print(merged_inner)
```
```{r }
not_merged <- anti_join(merged_left, merged_right)
print(not_merged)
```
